Live Functional Programming with Typed Holes

### 2018-02-26 calculus.tex
multi step, hole instantiation, resumption

\input{fig-substitution}

@@

\begin{theorem}[Commutativity \cy{state this}]
\end{theorem}

@@

\cy{canonical forms}

\cy{indeterminate forms}

>>>>

\input{fig-substitution}
\input{fig-multi-step}

@@

\rkc{some of these statements may need additional typing premises}

\begin{theorem}[Instantiation]
  If $\hasType{\hDelta}{\hGamma}{\dexp}{\tau}$
  and $\Dbinding{u}{\hGamma'}{\htau'}\in\hDelta$
  and $\hasType{\hDelta}{\hGamma'}{\dexp'}{\htau'}$
  then $\hasType{\hDelta}{\hGamma}{\instantiate{\dexp'}{u}{\dexp}}{\tau}$.
\end{theorem}

\autoref{fig:multi-step} defines the multi-step relation as the reflexive,
transitive closure of the single-step relation.

\begin{lemma}[Finality]
  If $\isFinal{\dexp}$ and $\multiStepsTo{\dexp}{\dexp'}$
  then $\dexp=\dexp'$.
\end{lemma}

\begin{theorem}[Commutativity]
  If $\multiStepsTo{\dexp_0}{\dexp_1}$
  then $\multiStepsTo{\instantiate{\dexp}{u}{\dexp_0}}
                     {\instantiate{\dexp}{u}{\dexp_1}}$.
\end{theorem}

\begin{theorem}[Confluence]
  If $\multiStepsTo{\dexp}{\dexp_1}$
  and $\multiStepsTo{\dexp}{\dexp_2}$
  then there exists $\dexp'$ such that
  $\multiStepsTo{\dexp_1}{\dexp'}$ and
  $\multiStepsTo{\dexp_2}{\dexp'}$.
\end{theorem}

\begin{corollary}[Final Confluence]
  If $\multiStepsTo{\dexp}{\dexp_1}$
  and $\isFinal{\dexp_1}$
  and $\multiStepsTo{\dexp}{\dexp_2}$
  then $\multiStepsTo{\dexp_2}{\dexp_1}$.
  \begin{proof}
    By Confluence and Finality.
  \end{proof}
\end{corollary}

\begin{theorem}[Resumption]
  If $\multiStepsTo{\dexp_1}{\dexp_2}$
  and $\isFinal{\dexp_2}$
  and $\multiStepsTo{\instantiate{\dexp_3}{u}{\dexp_1}}{\dexp_4}$
  and $\isFinal{\dexp_4}$
  then $\multiStepsTo{\instantiate{\dexp_3}{u}{\dexp_2}}{\dexp_4}$.
  \begin{proof}
    By Commutativity,
    $\multiStepsTo{\instantiate{\dexp_3}{u}{\dexp_1}}
                  {\instantiate{\dexp_3}{u}{\dexp_2}}$.
    By Final Confluence, we can conclude.
  \end{proof}
\end{theorem}

\begin{comment}

@@

\end{comment}
